Title Goes Here

### 2020-05-15 analysis/paper/paper.Rmd
some more background

Gender under-representation and inequality has long existed in science, from under-representation of women and minority genders in school text books to academic publications (@hamilton2006gender; @tushingham2017peer). Women are often under-represented for several reasons, including stereotyping that unfairly underestimates individuals' abilities, which often leads to lack of support for women in academic fields (@xu2008gender). In this paper we explore the relationship between gender and research topics to determine if women and men tend to focus more on certain topics than others.

@@

When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field (@bardolph2014critical).

As we can see in the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men (@heath2020guest).

When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership and the gender gap is almost gone. In this study we investigate gender ratios of presenters at major archaeological conferences.

@@

We estimated the gender of the first-named speaker for each presentation using the R programming language and the gender package (@lincolngender; @mihaljevic2019reflections; @blevins2015jane). We inferred the gender of personal names by looking each name up in the US Social Security Administration (SSA) baby name data, and calculating the overall probability that a given name was male or female (@mihaljevic2019reflections). If the proportion of people in the SSA data with a certain given name are recorded as female is 0.5 or higher, we predicted the gender of the presenter with this name as female. For example, in the SSA dataset the the name "Lynne" returns `r lynne_gender$proportion_male` as proportion of individuals with this name who are recorded as male, and `r lynne_gender$proportion_female` as the proportion female, thus we inferred that people in our conference data who are called Lynne are women.

@@

A structural topic model (STM) allows us to include, alongside the typical topic modeling process, other document-level metadata to analyse covariance between topics and metadata (@roberts2019package). Covariates of interest can be included into the prior distributions for document-topic proportions and
topic-word distributions using a standard regression model (cite this one: http://scholar.harvard.edu/files/dtingley/files/topicmodelsopenendedexperiments.pdf). This means that we can examine the relationship between the topics and document-level variables of interest. In our case, we computed the relationship between the inferred the gender of the first-named presenter for each presentation, and the distribution of topics in each presentation. When we observed the topics for each presentation, we generated a regression where the topic is the outcome variable, and gender is the explanatory variable. This regression gives us insights into whether gender caused archaeologists to spend a larger portion of their presentation on a particular topic.

>>>>

Gender under-representation and inequality has long existed in science, from under-representation of women and minority genders in school text books to academic publications [@hamilton2006gender; @tushingham2017peer]. Women are often under-represented for several reasons, including stereotyping that unfairly underestimates individuals' abilities, which often leads to lack of support for women in academic fields [@xu2008gender]. In this paper we explore the relationship between gender and research topics to determine if women and men tend to focus more on certain topics than others.

@@

When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field [@bardolph2014critical]. In a previous study done on gender balance of the production of the Atapuerca Project, a project about human evolution in the field of archaeo-palaeontology, they have found that among all the authors, only 35.9% of them are women [@rodriguez2018gender]. In our study, we also continue the study of gender distribution of presenters at major archaeological conferences.

Another similar imbalance can be seen in terms of the authorship of articles or papers written. For example, for the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men. [@heath2020guest]. This can also be seen in the study of the Atapuerca Project, in which only 34.8% of the papers are written by a female first/lead author [@rodriguez2018gender]. We have extended the study of ratios of authorships in the field of archaeology by specifically focusing on the gender imbalance that occurs in the topics that the authors present in major archaeological conferences.

When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership and the gender gap is almost gone. Although this similar trend is also present in the study by @vanderwarker2018ucsb in terms of more and more women pursuing a degree in archaeology than men, it is important to note that data collected from University of California, Santa Barbara Gender Equity Project shows that there are still imbalance as a result of different pay rates, professional ranking, and professional advancement [@gonzalez2018perceptions].

@@

We estimated the gender of the first-named speaker for each presentation using the R programming language and the gender package [@lincolngender; @mihaljevic2019reflections; @blevins2015jane]. We inferred the gender of personal names by looking each name up in the US Social Security Administration (SSA) baby name data, and calculating the overall probability that a given name was male or female [@mihaljevic2019reflections]. If the proportion of people in the SSA data with a certain given name are recorded as female is 0.5 or higher, we predicted the gender of the presenter with this name as female. For example, in the SSA dataset the the name "Lynne" returns `r lynne_gender$proportion_male` as proportion of individuals with this name who are recorded as male, and `r lynne_gender$proportion_female` as the proportion female, thus we inferred that people in our conference data who are called Lynne are women.

@@

A structural topic model (STM) allows us to include, alongside the typical topic modeling process, other document-level metadata to analyse covariance between topics and metadata [@roberts2019package]. Covariates of interest can be included into the prior distributions for document-topic proportions and
topic-word distributions using a standard regression model (cite this one: http://scholar.harvard.edu/files/dtingley/files/topicmodelsopenendedexperiments.pdf). This means that we can examine the relationship between the topics and document-level variables of interest. In our case, we computed the relationship between the inferred the gender of the first-named presenter for each presentation, and the distribution of topics in each presentation. When we observed the topics for each presentation, we generated a regression where the topic is the outcome variable, and gender is the explanatory variable. This regression gives us insights into whether gender caused archaeologists to spend a larger portion of their presentation on a particular topic.
